---
title: "Soil and Litter GitHub Issues Tracker"
author:
  name: Lai, Hao Ran
date: last-modified
format:
  html:
    other-links:
      - text: "GitHub Issues"
        icon: github
        href: https://github.com/ImperialCollegeLondon/ve_data_science/issues?q=is%3Aissue%20state%3Aopen%20author%3Ahrlai
      - text: "GitHub Milestones"
        icon: github
        href: https://github.com/ImperialCollegeLondon/ve_data_science/milestones
title-block-banner: true
comments:
  hypothesis: true
execute:
  echo: false
  warning: false
---

```{r}
#| label: setup
library(DT)
library(tidyverse)
library(IssueTrackeR)
library(htmltools)
library(here)

# IssueTrackerR configs
options(IssueTrackeR.dataset.dir = here::here("data/gh"))
options(IssueTrackeR.owner = "ImperialCollegeLondon")
options(IssueTrackeR.repo = "ve_data_science")

# Set global options for all DT tables
options(
  DT.options = list(
    pageLength = 5,
    autoWidth = TRUE,
    # 2. Pre-populate specific column filters
    searchCols = list(
      NULL,
      NULL,
      NULL,
      list(search = '["open"]'),
      NULL
    )
  )
)

# function to render the interactive table
render_issues <- function(issue) {
  issue |>
    select(-milestone) |>
    datatable(
      rownames = FALSE,
      colnames = c(
        "Issue" = "number",
        "Title" = "title",
        "Body (truncated)" = "body",
        "State" = "state",
        "Created at" = "created_at"
      ),
      filter = "top",
      escape = 2
    )
}

# function for a HTML progress bar
progress_bar <- function(x, total = 100, color = "#2c7bb6") {
  pct <- round(x / total * 100)
  lbl <- paste0(pct, "% complete; ", total - x, " open; ", x, " closed")

  tags$div(
    style = "margin: 1rem 0",
    tags$div(
      style = "display: flex; justify-content: space-between; font-size: 0.85rem; color: #555; margin-bottom: 4px;",
      tags$span("Progress"),
      tags$span(lbl)
    ),
    tags$div(
      style = "background: #e9ecef; border-radius: 999px; height: 10px; overflow: hidden;",
      tags$div(
        style = glue::glue(
          "width: {pct}%; height: 100%; background: {color};
           border-radius: 999px; transition: width 0.6s ease;"
        )
      )
    )
  )
}

# function that applies progress_bar() to a milestone object
progress_bar_milestone <- function(milestone, ...) {
  tmp <- milestones |> filter(title == milestone)
  n_closed <- tmp$nb_issues_closed
  n_total <- tmp$nb_issues_open + tmp$nb_issues_closed
  progress_bar(n_closed, n_total, ...)
}
```

```{r}
#| label: read-issues
# download GitHub issues
issues <- get_issues(source = "online", state = "all", verbose = FALSE)
milestones <- get_milestones(source = "online", state = "all", verbose = FALSE)

# write_to_dataset(issues)
# update_database()

# prepare the issue tracker table
issue_table <-
  issues |>
  as_tibble() |>
  filter(creator == "hrlai") |>
  # Convert NA milestones to a string for grouping
  mutate(milestone = replace_na(as.character(milestone), "No Milestone")) |>
  # Arrange by milestone, then by our custom state order
  mutate(state = factor(state, levels = c("open", "closed"))) |>
  arrange(milestone, state, desc(created_at)) |>
  # cleaner dates
  mutate_at(vars(created_at, closed_at), ymd) |>
  # Create a column for the linked number
  mutate(
    number = sprintf('<a href="%s" target="_blank">%s</a>', html_url, number),
    body = str_trunc(body, 100),
  ) |>
  select(number, title, body, state, milestone, created_at)
```

# Overview {#sec-overview}

## Progress: closed Issues and PRs
I updated the validation workflow by refactoring dataset screening ([#514](https://github.com/ImperialCollegeLondon/ve_data_science/pull/514 "Validation: Refactor log_dataset to screen_dataset and add_schema to improve dataset screening")), switching to a per-dataset schema ([#519](https://github.com/ImperialCollegeLondon/ve_data_science/pull/519 "Validation: switch to per-dataset schema for improved maintainability")), refining database building ([#522](https://github.com/ImperialCollegeLondon/ve_data_science/pull/522 "Validation: refine build_validation_database()")), and triaging soil-litter issues ([#531](https://github.com/ImperialCollegeLondon/ve_data_science/issues/531 "Triage soil-litter Issues to improve progress reporting")). I also completed Stage 2 of adding validation datasets ([#525](https://github.com/ImperialCollegeLondon/ve_data_science/issues/525 "Add soil and litter validation datasets (stage 2)")) and screened more soil and litter datasets ([#526](https://github.com/ImperialCollegeLondon/ve_data_science/pull/526 "Validation: screen more soil and litter datasets")). Codebase and tooling updates included adapting `subset_nc()` to handle Zarr outputs ([#510](https://github.com/ImperialCollegeLondon/ve_data_science/pull/510 "509 update subset_nc()"), [#509](https://github.com/ImperialCollegeLondon/ve_data_science/issues/509 "Update subset_nc() to cope with Zarr output vs netCDF input")), parsing the VE codebase to build a constants database for calibration ([#508](https://github.com/ImperialCollegeLondon/ve_data_science/pull/508 "Return of the Jedi: parsing VE codebase to build a constant database")), documenting sourced functions in R templates ([#292](https://github.com/ImperialCollegeLondon/ve_data_science/issues/292 "Document functions that are sourced from another script")), and explicitly referencing external R function dependencies ([#297](https://github.com/ImperialCollegeLondon/ve_data_science/issues/297 "Explicitly reference dependencies of external functions in tools/R")). I closed the enzyme concentration modeling issue because it will not be addressed in the near term ([#63](https://github.com/ImperialCollegeLondon/ve_data_science/issues/63 "Model enzyme / functional gene concentrations / compositions")).

## TODO: pending Issues and PRs
### Urgent and High priority
Combining the newly added soil gas flux validation dataset with VE outputs is currently blocked, awaiting VE updates ([#62](https://github.com/ImperialCollegeLondon/ve_data_science/issues/62 "Validate soil gas fluxes")). I also need to recompile soil input data ([#538](https://github.com/ImperialCollegeLondon/ve_data_science/pull/538 "489 recompile soil input data")).

### Medium and Low priority
I need to update the LLM prompt to group soil microbial and enzyme constants ([#534](https://github.com/ImperialCollegeLondon/ve_data_science/issues/534 "LLM: update prompt to treat SoilMicrobialGroup or SoilEnzymeClass constants as a group")), implement RAG for VE docs after reverting changes in #506 ([#532](https://github.com/ImperialCollegeLondon/ve_data_science/issues/532 "LLM: RAG for VE docs")), and add a raw dataset filtering option to `build_validation_database` ([#523](https://github.com/ImperialCollegeLondon/ve_data_science/issues/523 "build_validation_database needs a filtering option for raw datasets")).

# Milestones {#sec-milestones}

## Soil parameterisation

Parameterise soil and litter constants with empirical data or best-guess estimates.

```{r}
#| label: parameterisation
# milestone overview
progress_bar_milestone("Soil parameterisation")

# issue table
issue_table |>
  filter(milestone == "Soil parameterisation") |>
  render_issues()
```

## Soil validation

Validate soil and litter variables.

```{r}
#| label: validation
# milestone overview
progress_bar_milestone("Soil validation")

# issue table
issue_table |>
  filter(milestone == "Soil validation") |>
  render_issues()
```

## Soil calibration

Not started yet.

## Soil sensitivity

Sensitivity analysis for soil and litter.

```{r}
#| label: sensitivity
# milestone overview
progress_bar_milestone("Soil sensitivity")

# issue table
issue_table |>
  filter(milestone == "Soil sensitivity") |>
  render_issues()
```

## Soil input data (phase 2)

Follow up to polish soil and litter input data for completeness.

```{r}
#| label: input2
# milestone overview
progress_bar_milestone("Soil input data (phase 2)")

# issue table
issue_table |>
  filter(milestone == "Soil input data (phase 2)") |>
  render_issues()
```

## Soil input data (phase 1)

MVP of soil and litter input datasets using "Maliau" datasets. This milestone has been fully completed.

```{r}
#| label: input1
# milestone overview
progress_bar_milestone("Soil input data (phase 1)")

# issue table
issue_table |>
  filter(milestone == "Soil input data (phase 1)") |>
  render_issues()
```

## Other tasks

These are miscellaneous tasks not targetting any milestones, but are important for housekeeping and repo maintenance.

```{r}
#| label: other-tasks
# issue table
issue_table |>
  filter(milestone == "No Milestone") |>
  render_issues()
```
