---
title: "Soil and Litter GitHub Issues Tracker"
author:
  name: Lai, Hao Ran
date: last-modified
format:
  html:
    other-links:
      - text: "GitHub Issues"
        icon: github
        href: https://github.com/ImperialCollegeLondon/ve_data_science/issues?q=is%3Aissue%20state%3Aopen%20author%3Ahrlai
      - text: "GitHub Milestones"
        icon: github
        href: https://github.com/ImperialCollegeLondon/ve_data_science/milestones
title-block-banner: true
comments:
  hypothesis: true
execute:
  echo: false
  warning: false
---

```{r}
#| label: setup
library(DT)
library(tidyverse)
library(IssueTrackeR)
library(htmltools)
library(here)

# IssueTrackerR configs
options(IssueTrackeR.dataset.dir = here::here("data/gh"))
options(IssueTrackeR.owner = "ImperialCollegeLondon")
options(IssueTrackeR.repo = "ve_data_science")

# Set global options for all DT tables
options(
  DT.options = list(
    pageLength = 5,
    autoWidth = TRUE,
    # 2. Pre-populate specific column filters
    searchCols = list(
      NULL,
      NULL,
      NULL,
      list(search = '["open"]'),
      NULL
    )
  )
)

# function to render the interactive table
render_issues <- function(issue) {
  issue |>
    select(-milestone) |>
    datatable(
      rownames = FALSE,
      colnames = c(
        "Issue" = "number",
        "Title" = "title",
        "Body (truncated)" = "body",
        "State" = "state",
        "Created at" = "created_at"
      ),
      filter = "top",
      escape = 2
    )
}

# function for a HTML progress bar
progress_bar <- function(x, total = 100, color = "#2c7bb6") {
  pct <- round(x / total * 100)
  lbl <- paste0(pct, "% complete; ", total - x, " open; ", x, " closed")

  tags$div(
    style = "margin: 1rem 0",
    tags$div(
      style = "display: flex; justify-content: space-between; font-size: 0.85rem; color: #555; margin-bottom: 4px;",
      tags$span("Progress"),
      tags$span(lbl)
    ),
    tags$div(
      style = "background: #e9ecef; border-radius: 999px; height: 10px; overflow: hidden;",
      tags$div(
        style = glue::glue(
          "width: {pct}%; height: 100%; background: {color};
           border-radius: 999px; transition: width 0.6s ease;"
        )
      )
    )
  )
}

# function that applies progress_bar() to a milestone object
progress_bar_milestone <- function(milestone, ...) {
  tmp <- milestones |> filter(title == milestone)
  n_closed <- tmp$nb_issues_closed
  n_total <- tmp$nb_issues_open + tmp$nb_issues_closed
  progress_bar(n_closed, n_total, ...)
}
```

```{r}
#| label: read-issues
# download GitHub issues
issues <- get_issues(source = "online", state = "all", verbose = FALSE)
milestones <- get_milestones(source = "online", state = "all", verbose = FALSE)

# write_to_dataset(issues)
# update_database()

# prepare the issue tracker table
issue_table <-
  issues |>
  as_tibble() |>
  filter(creator == "hrlai") |>
  # Convert NA milestones to a string for grouping
  mutate(milestone = replace_na(as.character(milestone), "No Milestone")) |>
  # Arrange by milestone, then by our custom state order
  mutate(state = factor(state, levels = c("open", "closed"))) |>
  arrange(milestone, state, desc(created_at)) |>
  # cleaner dates
  mutate_at(vars(created_at, closed_at), ymd) |>
  # Create a column for the linked number
  mutate(
    number = sprintf('<a href="%s" target="_blank">%s</a>', html_url, number),
    body = str_trunc(body, 100),
  ) |>
  select(number, title, body, state, milestone, created_at)
```

# Overview {#sec-overview}

## Urgent and High priority Issues

*   **Validation Database Refactoring**: We updated our validation pipeline by transitioning to a per-dataset schema ([#519](https://github.com/ImperialCollegeLondon/ve_data_science/pull/519 "Validation: switch to per-dataset schema for improved maintainability")), refactoring screening logs ([#514](https://github.com/ImperialCollegeLondon/ve_data_science/pull/514 "Validation: Refactor log_dataset to screen_dataset and add_schema to improve dataset screening")), and refining the database build process ([#522](https://github.com/ImperialCollegeLondon/ve_data_science/pull/522 "Validation: refine build_validation_database()")). 
*   **Soil and Litter Datasets**: We screened more soil and litter datasets ([#526](https://github.com/ImperialCollegeLondon/ve_data_science/pull/526 "Validation: screen more soil and litter datasets")) to address [#62](https://github.com/ImperialCollegeLondon/ve_data_science/issues/62 "Issue 62"), though combining this with Virtual Ecosystem outputs is currently pending external repository updates. We also opened PR [#538](https://github.com/ImperialCollegeLondon/ve_data_science/pull/538 "489 recompile soil input data") to recompile soil input data, and are coordinating a revert in [#532](https://github.com/ImperialCollegeLondon/ve_data_science/issues/532 "Issue 532") before proceeding.

## Medium and Low priority Issues

*   **Tooling and Core Codebase**: We integrated a Shiny dashboard ([#520](https://github.com/ImperialCollegeLondon/ve_data_science/pull/520 "Validation: Shiny dashboard for dataset screening")) as a new workflow tool to streamline dataset screening. Additionally, we parsed the Virtual Ecosystem codebase to build a constant database ([#508](https://github.com/ImperialCollegeLondon/ve_data_science/pull/508 "Return of the Jedi: parsing VE codebase to build a constant database")) and updated the netCDF subsetting function ([#510](https://github.com/ImperialCollegeLondon/ve_data_science/pull/510 "509 update subset_nc()")).
*   **Issue Cleanup and Discussions**: We closed several resolved or stale issues, including [#292](https://github.com/ImperialCollegeLondon/ve_data_science/issues/292 "Issue 292") (R template script field resolution), [#297](https://github.com/ImperialCollegeLondon/ve_data_science/issues/297 "Issue 297") (R namespace verification), and [#63](https://github.com/ImperialCollegeLondon/ve_data_science/issues/63 "Issue 63") (deferred). Discussions remain active regarding notebook directory layouts ([#515](https://github.com/ImperialCollegeLondon/ve_data_science/issues/515 "Issue 515")), DOI/URL metadata fields ([#433](https://github.com/ImperialCollegeLondon/ve_data_science/issues/433 "Issue 433")), and literature search queries for Maliau Basin ([#525](https://github.com/ImperialCollegeLondon/ve_data_science/issues/525 "Issue 525")).

## Pending PRs

- Urgent
  - TaranRallings, nickwctan, Baizurah to review PR [#330](https://github.com/ImperialCollegeLondon/ve_data_science/pull/330) — *329 examine animal continuous state variables from `maliau_2`*
  - nickwctan, arne-exe to review PR [#538](https://github.com/ImperialCollegeLondon/ve_data_science/pull/538) — *489 recompile soil input data*
- Not urgent
  - annarallings, davidorme to review PR [#422](https://github.com/ImperialCollegeLondon/ve_data_science/pull/422) — *Pin mkdocs-material version to 9.7.0*

## Additional notes

# Subtasks (as GitHub Milestones) {#sec-milestones}

## Soil parameterisation

Parameterise soil and litter constants with empirical data or best-guess estimates.

```{r}
#| label: parameterisation
# milestone overview
progress_bar_milestone("Soil parameterisation")

# issue table
issue_table |>
  filter(milestone == "Soil parameterisation") |>
  render_issues()
```

## Soil validation

Validate soil and litter variables.

```{r}
#| label: validation
# milestone overview
progress_bar_milestone("Soil validation")

# issue table
issue_table |>
  filter(milestone == "Soil validation") |>
  render_issues()
```

## Soil calibration

Not started yet.

## Soil sensitivity

Sensitivity analysis for soil and litter.

```{r}
#| label: sensitivity
# milestone overview
progress_bar_milestone("Soil sensitivity")

# issue table
issue_table |>
  filter(milestone == "Soil sensitivity") |>
  render_issues()
```

## Soil input data (phase 2)

Follow up to polish soil and litter input data for completeness.

```{r}
#| label: input2
# milestone overview
progress_bar_milestone("Soil input data (phase 2)")

# issue table
issue_table |>
  filter(milestone == "Soil input data (phase 2)") |>
  render_issues()
```

## Soil input data (phase 1)

MVP of soil and litter input datasets using "Maliau" datasets. This milestone has been fully completed.

```{r}
#| label: input1
# milestone overview
progress_bar_milestone("Soil input data (phase 1)")

# issue table
issue_table |>
  filter(milestone == "Soil input data (phase 1)") |>
  render_issues()
```

## Other tasks

These are miscellaneous tasks not targetting any milestones, but are important for housekeeping and repo maintenance.

```{r}
#| label: other-tasks
# issue table
issue_table |>
  filter(milestone == "No Milestone") |>
  render_issues()
```
